Manuscript progress report

### 2024-05-28 Docs/manuscript.Rmd
presentation papers

```{r fig_large_sim, fig.height=8, fig.width=8}
bind_rows(
  pred_list_all[[1]] %>% filter(id %in% rand_id) %>% 
    mutate(method="fGFPCA"),
  pred_list_ref[[1]] %>% filter(id %in% rand_id) %>%
    mutate(method = "GLMMadaptive"),
  pred_list_gfosr_l1[[1]] %>% filter(id %in% rand_id) %>%
    mutate(method = "GFOSR (L=1)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_list_gfosr_l5[[1]] %>% filter(id %in% rand_id) %>%
    mutate(method = "GFOSR (L=5)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4)
) %>% mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), linetype = "dashed")+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), na.rm=T, linewidth = 1.0)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(legend.position = "bottom")+
  scale_color_manual(values = cols)
```

@@

## Time (minutes)

I chose not to report fitting and prediction time separately because it is too difficult to tell them apart for the GFOSR models.

```{r}
data.frame(
  "Method" = c("fGFPCA", "GLMMadaptive", "GFOSR (L=1)", "GFOSR (L=5)"),
  "Time" = c(mean(fit_time/60+pred_time),
             mean(fit_time_ref+pred_time_ref/60),
             mean(t_vec_gfosr_l1/60), 
             mean(t_vec_gfosr_l5/60))) %>%
  kable(digits = 3, booktabs = T, table.attr = "style = \"color:black;\"") %>%
  kable_styling(full_width = F) 
```

@@

```{r fig_small_sim, fig.height=8, fig.width=8}
bind_rows(
  pred_subset_fGFPCA[[1]] %>% filter(id %in% rand_id) %>% mutate(method="fGFPCA"),
  pred_subset_gfosr_l5[[1]] %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=5)") %>% 
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_gfosr_l1[[1]] %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=1)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_adglmm[[1]] %>% filter(id %in% rand_id) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), linetype = "dashed")+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linewidth=1, na.rm=T)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(legend.position = "bottom")+
  scale_color_manual(values = cols)
```

@@

```{r}
data.frame(
  "Method" = c("fGFPCA", "GLMMadaptive", "GFOSR (L=1)", "GFOSR (L=5)"),
  "Time" = c(mean(fit_time_subset_fGFPCA/60+pred_time_subset_fGFPCA),
             mean(fit_time_subset_adglmm+pred_time_subset_adglmm/60),
             mean(t_subset_gfosr_l1/60), 
             mean(t_subset_gfosr_l5/60))) %>%
  kable(digits = 3, booktabs = T, table.attr = "style = \"color:black;\"") %>%
  kable_styling(full_width = F) 
```

@@

```{r fig_appl, fig.height=8, fig.width=8}
# redefine color scale
cols <- cols[1:3] 
names(cols) <- c("6am", "12pm", "6pm")

@@

@@

# Additional figures

```{r eval=FALSE}
df_nhanes <- read_rds(here("Data/nhanes_bi.rds"))
df_nhanes %>% 
  select(SEQN, Z, sind) %>%
  filter(SEQN %in% rand_id) %>%
  filter(sind %in% 356:360) %>%
  pivot_wider(names_from = sind, values_from = Z) %>% 
  select(-SEQN) %>% distinct(.)

pred_appl_gfosr_l5 %>% 
  filter(id %in% rand_id) %>%
  select(id, sind, pred360) %>%
  pivot_wider(names_from = sind, values_from = pred360) %>% 
  select(-id) %>% distinct(.)
```

<!-- ## fGFPCA algorithm demonstration -->

<!-- ```{r binning, eval=FALSE} -->
<!-- df_nhanes %>% select(SEQN, Z, sind) %>%  -->
<!--   filter(SEQN %in% rand_id3 & sind >= 650 & sind <= 750) %>%  -->
<!--   mutate(SEQN = as.factor(SEQN)) %>%  -->
<!--   ggplot()+ -->
<!--   geom_point(aes(x=sind, y = Z, col = SEQN), size = 0.5)+ -->
<!--   geom_vline(xintercept = seq(650, 750, by = 10), linetype="dashed")+ -->
<!--   scale_y_continuous(breaks = c(0, 1))+ -->
<!--   labs(x="Time", y="", title = "NHANES binary activity indicator") -->
<!-- ``` -->

<!-- ```{r,eval=FALSE} -->
<!-- df_fig <- df_nhanes %>% select(SEQN, Z, sind) %>%  -->
<!--   filter(SEQN %in% rand_id3 & sind >= 650 & sind <= 750) %>%  -->
<!--   mutate(bin = cut(sind, breaks = seq(650, 750, by = 10),  -->
<!--                    labels = seq(655, 745, by = 10), -->
<!--                    include.lowest = T))  -->
<!-- bin_mid <- seq(655, 745, by = 10) -->

<!-- for(i in bin_mid){ -->
<!--   df_bin_i <- df_fig %>% filter(bin==i)  -->
<!--   glm_i <- glmer(Z ~ 1 + (1|SEQN), data = df_bin_i, family = binomial, nAGQ=0) -->
<!--   eta_i <- predict(glm_i, type = "link") -->

<!--   df_fig$eta_hat[df_fig$bin==i] <- eta_i -->
<!-- } -->

<!-- df_fig %>%  -->
<!--   mutate(SEQN = as.factor(SEQN), -->
<!--          eta_hat = exp(eta_hat)/(1+exp(eta_hat)), -->
<!--          bin=as.numeric(as.character(bin))) %>%  -->

<!--   ggplot()+ -->
<!--   geom_point(aes(x=sind, y = Z, col = SEQN), size = 0.5)+ -->
<!--   geom_vline(xintercept = seq(650, 750, by = 10), linetype="dashed")+ -->
<!--   geom_point(aes(x=bin, y=eta_hat, col = SEQN))+ -->
<!--   geom_line(aes(x=bin, y=eta_hat, col = SEQN))+ -->
<!--   scale_y_continuous(breaks = c(0, 1))+ -->
<!--   labs(x="Time", y="", title = "NHANES binary activity indicator") -->

<!-- ``` -->

<!-- ## Example of continuous and binary NHANES -->

<!-- ```{r, eval=FALSE} -->
<!-- df_nhanes_cont <- readRDS(here("Data/NHANES_AC_processed.rds")) -->
<!-- p_cont <- df_nhanes_cont %>%  -->
<!--   filter(SEQN %in% sample(unique(df_nhanes_cont$SEQN), 4) & WEEKDAY==6) %>%  -->
<!--   select(starts_with("min"), SEQN) %>% -->
<!--   pivot_longer(1:1440, names_to = "Minute") %>%  -->
<!--   mutate(Minute = gsub("MIN", "", Minute)) %>% -->
<!--   mutate(Minute = as.numeric(Minute), -->
<!--          SEQN=as.factor(SEQN)) %>% -->
<!--   ggplot(aes(x=Minute, y=value, col=SEQN))+ -->
<!--   geom_point(size = 0.5, alpha=0.5)+ -->
<!--   geom_smooth(se=F, size = 0.5)+ -->
<!--   scale_x_continuous(breaks = seq(0, 1440, by = 480), -->
<!--                      labels = c("Midnight", "8am", "4pm", "Midnight"))+ -->
<!--   theme(axis.text = element_text(size = 5))+ -->
<!--   labs(y="", title = "NHANES Activitiy count (2003-2004)") -->
<!-- ``` -->

<!-- ```{r, eval=FALSE} -->
<!-- # a example of continuous and binary NHANES data -->
<!-- p_bi <- df_nhanes %>% select(SEQN, Z, sind) %>%  -->
<!--   filter(SEQN %in% rand_id3 & Z == 1) %>%  -->
<!--   mutate(SEQN = as.factor(SEQN)) %>%  -->
<!--   ggplot()+ -->
<!--   geom_point(aes(x=sind, y = SEQN), size = 0.5)+ -->
<!--   scale_x_continuous(breaks = seq(0, 1440, by = 480), -->
<!--                      labels = c("Midnight", "8am", "4pm", "Midnight"))+ -->
<!--   theme(axis.text = element_text(size = 5))+ -->
<!--   labs(x="Time", y="Subject", title = "NHANES binary activity indicator (2011-2014)") -->
<!-- ``` -->

<!-- ```{r data_exp, fig.height=4, fig.width=10, eval=FALSE} -->
<!-- ggarrange(p_cont, p_bi, nrow = 1) -->
<!-- ``` -->

<!-- ## Dynamic prediction example -->

<!-- ```{r dynpred_exp,  fig.height=4, fig.width=10, eval=FALSE} -->
<!-- ggarrange( -->
<!--   pred_list_all[[1]] %>%  -->
<!--     filter(id==501) %>% -->
<!--     mutate(eta_i = ifelse(t<=0.2, eta_i, NA)) %>% -->
<!--     ggplot()+ -->
<!--     geom_line(aes(x=t, y=eta_i, col = "Observed"), na.rm = T)+ -->
<!--     geom_line(aes(x=t, y=pred0.2, col = "Predicted"), na.rm = T,  -->
<!--               linetype = "dashed")+ -->
<!--     geom_vline(xintercept = 0.2, linetype = "dashed")+ -->
<!--     labs(x="Time", y="")+ -->
<!--     scale_x_continuous(breaks = seq(0, 1, by = 0.2))+ -->
<!--     theme(axis.text = element_text(size = 5))+ -->
<!--     ylim(-1.2, 1.7), -->

<!--  pred_list_all[[1]] %>%  -->
<!--     filter(id==501) %>% -->
<!--     mutate(eta_i = ifelse(t<=0.4, eta_i, NA)) %>% -->
<!--     ggplot()+ -->
<!--     geom_line(aes(x=t, y=eta_i, col = "Observed"), na.rm = T)+ -->
<!--     geom_line(aes(x=t, y=pred0.4, col = "Predicted"), na.rm = T,  -->
<!--               linetype = "dashed")+ -->
<!--     geom_vline(xintercept = 0.4, linetype = "dashed")+ -->
<!--     labs(x="Time", y="")+ -->
<!--     scale_x_continuous(breaks = seq(0, 1, by = 0.2))+ -->
<!--     theme(axis.text = element_text(size = 5))+ -->
<!--    ylim(-1.2, 1.7), -->
<!--  nrow = 1, common.legend = T -->
<!--   ) -->

<!-- ``` -->

>>>>

```{r fig_large_sim_paper, fig.height=8, fig.width=8}
bind_rows(
  pred_list_all[[1]] %>% filter(id %in% rand_id) %>% 
    mutate(method="fGFPCA"),
  pred_list_ref[[1]] %>% filter(id %in% rand_id) %>%
    mutate(method = "GLMMadaptive"),
  pred_list_gfosr_l1[[1]] %>% filter(id %in% rand_id) %>%
    mutate(method = "GFOSR (L=1)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_list_gfosr_l5[[1]] %>% filter(id %in% rand_id) %>%
    mutate(method = "GFOSR (L=5)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4)
) %>% mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), linetype = "dashed")+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), na.rm=T, linewidth = 1.0)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), na.rm=T, linewidth = 1.0)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(legend.position = "bottom")+
  scale_color_manual(values = cols)
```

@@

@@

```{r fig_small_sim_paper, fig.height=8, fig.width=8}
bind_rows(
  pred_subset_fGFPCA[[1]] %>% filter(id %in% rand_id) %>% mutate(method="fGFPCA"),
  pred_subset_gfosr_l5[[1]] %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=5)") %>% 
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_gfosr_l1[[1]] %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=1)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_adglmm[[1]] %>% filter(id %in% rand_id) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), linetype = "dashed")+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linewidth=1, na.rm=T)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))+
  theme(legend.position = "bottom")+
  scale_color_manual(values = cols)
```

```{r}
sep_plot <- bind_rows(
  pred_subset_fGFPCA[[1]] %>% filter(id==189) %>% mutate(method="fGFPCA"),
  pred_subset_gfosr_l5[[1]] %>% filter(id==189) %>% mutate(method = "GFOSR (L=5)") %>% 
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_gfosr_l1[[1]] %>% filter(id==189) %>% mutate(method = "GFOSR (L=1)") %>%
    rename(pred0.2=pred_w1, pred0.4=pred_w2, pred0.6=pred_w3, pred0.8=pred_w4),
  pred_subset_adglmm[[1]] %>% filter(id==189) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(eta_i, pred0.2, pred0.4, pred0.6, pred0.8), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=t, y=Y), size = 0.2)+
  geom_line(aes(x=t, y=eta_i, col = "True"), linetype = "dashed")+
  geom_line(aes(x=t, y=pred0.2, col = "0.2"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.4, col = "0.4"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.6, col = "0.6"), linewidth=1, na.rm=T)+
  geom_line(aes(x=t, y=pred0.8, col = "0.8"), linewidth=1, na.rm=T)+
  facet_grid(cols = vars(method))+
  labs(col = "Maximum observation time", x = "t", y="")+
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))
```

```{r, fig_small_sim_pre, fig.height=2, fig.width=8}
sep_plot+scale_color_manual(values = cols, guide="none")+
  #theme(legend.position = "bottom")+
  scale_color_manual(values = cols, guide = "none")
```

```{r, legend, fig.height=0.2, fig.width=8}
as_ggplot(ggpubr::get_legend(sep_plot+theme(legend.position = "bottom")+
                               scale_color_manual(values = cols)))
```

@@

Average time per iteration in minutes

```{r}
time_tb <- data.frame(sim1 = c(mean(fit_time/60+pred_time),
             mean(fit_time_ref+pred_time_ref/60),
             mean(t_vec_gfosr_l1/60), 
             mean(t_vec_gfosr_l5/60)),
           sim2 = c(mean(fit_time_subset_fGFPCA/60+pred_time_subset_fGFPCA),
             mean(fit_time_subset_adglmm+pred_time_subset_adglmm/60),
             mean(t_subset_gfosr_l1/60), 
             mean(t_subset_gfosr_l5/60))) %>% t()
colnames(time_tb) <- c("fGFPCA", "GLMMadaptive", "GFOSR (L=1)", "GFOSR (L=5)")
rownames(time_tb) <- c("Simulation 1", "Simulation 2")
```

```{r time_tb}
time_tb %>% 
  kable(digits = 3, booktabs = T, table.attr = "style = \"color:black;\"") %>%
  kable_styling(full_width = F) 
```

@@

```{r fig_appl_paper, fig.height=8, fig.width=8}
# redefine color scale
cols <- cols[1:3] 
names(cols) <- c("6am", "12pm", "6pm")

@@

```{r fig_appl_pre, fig.height=4, fig.width=8}
# selct four subjects
plot_id <- sort(c(64198, 80517))

@@
